---
title: "USA Dashboard of Novel Carona Virus Cases"
author: "updated"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
# Load Data 
library(flexdashboard)
library(tidyverse)
library(dcl)
library(lubridate)
library(leaflet)
library(leaflet.extras)
library(sf)
library(kableExtra)

#Parameters 

url_cases <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv"
url_deaths <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Deaths.csv"
url_recovered <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Recovered.csv"


#Rename Table 

vars_cases <- c("cases","deaths", "recovered")

vars_cases_recode <- 
  c(
    "active_cases" = "Active Cases",
    "recovered" = "Recovered",
    "cases" = "Cases Total", 
    "deaths" = "Deaths"
  )

vars_cases_color<- 
  c(
    "Active Cases" = "purple",
    "Recovered" = "dark green",
    "Cases Total" = "orange", 
    "Deaths" = "red"
  )

cases_colors <- 2

vars_totals <-
  c(
    `States Infected` = "t_states",
    `Total Cases` = "t_cases",
    `Total Deaths` = "t_deaths",
    `Total Recovered` = "t_recovered",
    `Mortality Rate %` = "av_mortality_rate"
  )

vars <-
  c(
    # `Countries` = "countries",
    `Total Cases` = "t_cases",
    `Total Deaths` = "t_deaths",
    `Total Recovered` = "t_recovered",
    `Active Cases` = "active_cases",
    `Mortality Rate %` = "av_mortality_rate"
  )

vars_rename <- 
  c(
    province = "Province/State",
    country = "Country/Region"
  )

country_recode <- 
  c(
    "Mainland China" = "China",
    # "Hong Kong" = "China", #included Hong Kong as part of China
    # "Taiwan" = "China",
    .default = "Other"
  )




```

```{r}
#Clean data

carona_datasets <- c(url_cases, url_deaths, url_recovered)

#Read Cases
cases_by_date <- 
  read_csv(url_cases) %>% 
  pivot_longer(
    cols = matches("^\\d"),
    names_to = "date",
    values_to = "cases"
  ) %>% 
  mutate(date = mdy(date))

# Read Deaths
deaths_by_date <- 
  read_csv(url_deaths) %>% 
  pivot_longer(
    cols = matches("^\\d"),
    names_to = "date",
    values_to = "deaths"
  ) %>% 
  mutate(date = mdy(date))

#Read Recovey
recovery_by_date <- 
  read_csv(url_recovered) %>% 
  pivot_longer(
    cols = matches("^\\d"),
    names_to = "date",
    values_to = "recovered"
  ) %>% 
  mutate(date = mdy(date))

#Join Tables
carona_data_usa <- 
  cases_by_date %>% 
  left_join(left_join(deaths_by_date, recovery_by_date)) %>% 
  rename(!!vars_rename) %>% 
  filter(country == "US") %>% 
  separate(province, c("city", "state"), sep = ", ") %>% 
  mutate(state = str_sub(state, start = 1, end = 3))

# carona_data_usa %>% 
#   filter(is.na(state))


carona_data_totals <- 
  carona_data_usa %>% 
  filter(date == max(date) & cases > 0)

```



Column {data-width=650}
-----------------------------------------------------------------------


### Total Cumulative Cases {data-height=150}

```{r}
carona_data_totals %>% 
  summarise(
    t_states = n_distinct(state), #-1 bc Hong Kong Seperate in dataset + Taiwan but i am calling Taiwan an independant country 
    t_cases = sum(cases),
    t_deaths = sum(deaths),
    av_mortality_rate = t_deaths / t_cases * 100,
    t_recovered = sum(recovered)
  ) %>% 
  # mutate(
  #   t_cases = call_spec(t_cases, colo)
  # ) %>% 
  rename(!!vars_totals) %>% 
  knitr::kable() %>% 
  kable_styling(font_size = 20) %>% 
  column_spec(2:3, bold = T)

#View(carona_data_totals)

# mutate(
#   car = row.names(.),
#   mpg = cell_spec(mpg, "html", color = ifelse(mpg > 20, "red", "blue")),
#   cyl = cell_spec(cyl, "html", color = "white", align = "c", angle = 45, 
#                   background = factor(cyl, c(4, 6, 8), 
#                                       c("#666666", "#999999", "#BBBBBB")))
# )

#### Try library GT for making pretty tables 
```
### Map of Cases By Country {data-height=650}
```{r}
#read in countries
#countries <- read_sf(file_location_countries)

# # Create a color palette with handmade bins.
# library(RColorBrewer)
# mybins <- c(0,1,5,10,50,100,500, Inf)
# mypalette <- colorBin( palette="YlOrBr", domain=carona_data_usa, na.color="transparent", bins=mybins)
#  
# # Prepare the text for tooltips:
# mytext <- paste(
#     "State: ", carona_data_usa$state,"<br/>", 
#     #"Area: ", world_spdf@data$AREA, "<br/>", 
#     "Cases: ", round(carona_data_usa$cases, 2), 
#     sep="") %>%
#   lapply(htmltools::HTML)
#  
# # Final Map
# leaflet() %>% 
#   addTiles()  %>% 
#   setView( lat=10, lng=0 , zoom=2) %>%
#   addPolygons( 
#     lng = Long,
#     lat = Lat,
#     fillColor = ~mypalette(cases), 
#     stroke=TRUE, 
#     fillOpacity = 0.9, 
#     color="white", 
#     weight=0.3,
#     label = mytext,
#     labelOptions = labelOptions( 
#       style = list("font-weight" = "normal", padding = "3px 8px"), 
#       textsize = "13px", 
#       direction = "auto"
#     )
#   ) %>%
#   addLegend( pal=mypalette, values=~cases, opacity=0.9, title = "Population (M)", position = "bottomleft" )
# 
# 
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircles(
    data = carona_data_totals %>% drop_na(state),
    # lng = Long,
    # lat = Lat,
    radius = ~ 1500 * sqrt(cases),
    color = "red",
    weight = 2,
    fillColor = "red",
    fillOpacity = 0.6,
    label =
      ~ str_c(
        if_else(!is.na(city), city, ""),
        ", ",
        if_else(!is.na(state), state, ""),
        " ", country,
        ": ", cases,
        if_else(cases == 1, " case", " cases")),

  )

# summary(carona_data$province)

```



### Cumulative Cases By Country {data-height=250}

```{r}
country_summary_data <- 
  carona_data_totals %>%  
  group_by(state) %>% 
  summarise(
    t_cases = sum(cases),
    t_deaths = sum(deaths),
    t_recovered = sum(recovered)
  ) %>% 
  mutate(
    active_cases = t_cases - t_deaths - t_recovered,
    av_mortality_rate = t_deaths / t_cases * 100
  ) %>% 
  arrange(desc(t_cases)) %>% 
  rename(!!vars)

country_summary_data %>% 
  knitr::kable() %>% 
  kable_styling(bootstrap_options = "striped", "responsive",full_width = F)


```

Column {.tabset .tabset-fade data-width=400}
-----------------------------------------------------------------------

### Cumulative Case Data

```{r, fig.width=5.5, fig.height=3.7}
#Cases by date tibble 
carona_data_by_date <- 
  carona_data_usa %>%
  group_by(date) %>% 
  summarise_at(vars(vars_cases), sum, na.rm = TRUE) %>%
  mutate(active_cases = cases - recovered - deaths) 

#Cases by date and country tibble 
# carona_data_by_country_date <- 
#   carona_data %>%
#   group_by(country_combined, date) %>% 
#   summarise_at(vars(vars_cases), sum, na.rm = TRUE) %>%
#   mutate(active_cases = cases - recovered - deaths) 


#Total cases by date plot
carona_data_by_date %>%
  pivot_longer(cols = c(cases, deaths, recovered, active_cases),
               names_to = "case_type",
               values_to = "number"
  ) %>%
  mutate(case_type = recode(case_type, !!! vars_cases_recode)) %>%  
  ggplot(aes(x = date, y = number, color = case_type)) +
  geom_point() +
  geom_line() + 
  ggrepel::geom_text_repel(
    aes(label = str_glue("{case_type}")), #Adding in {number} works but pushes this out of the table
    data = . %>% group_by(case_type) %>% top_n(n = 1, wt = date), #weight is variable to order by
    size = 3,
    hjust = -0.12,
    color = "black"
  ) +
  scale_x_date(
    date_breaks = "1 week",
    #date_breaks = seq(min(carona_data_by_date$date), max(carona_data_by_date$date), by = "1 week"),
    date_labels = "%b %d",
    expand = c(0, 2)
  ) +
  scale_y_continuous(
    breaks = scales::breaks_width(25),
    labels = scales::label_number(big.mark = ",")
  ) +
  scale_color_manual(values = vars_cases_color) +
  theme(legend.position = "none") +
  coord_cartesian(
    xlim = 
      c(
        min(carona_data_by_date$date),
        max(carona_data_by_date$date) + days(7)
      )
  ) + 
  labs(
    title = "Cumulative Number of Cases by Date",
    x = NULL, 
    y = "Number of Total Cases"
    #caption = "Source: John Hopkins COV-19 Database"
  )


# Total Cases by Date and country 

# carona_data_by_country_date %>% 
#   ggplot(aes(x = date, y = cases, color = state)) +
#   geom_point() +
#   geom_line() +
#   scale_x_date(
#     date_breaks = "1 week",
  #   date_labels = "%b %d",
  #   expand = c(0, 1)
  # ) +
  # scale_y_continuous(labels = scales::label_number(big.mark = ",")) +
  # theme(legend.position = "bottom") +
  # labs(
  #   title = "Total Number of Cases by Date",
  #   x = NULL, 
  #   y = "Number of Total Cases", 
  #   color = NULL,
  #   caption = "Source: John Hopkins COV-19 Database"
  # )

#New Cases by date 
carona_data_by_date %>% 
  mutate(
    new_cases = cases - lag(cases),
    new_recovered = recovered - lag(recovered)
  ) %>% 
  ggplot(aes(x = date)) +
  #geom_col(aes(y = new_recovered), fill = "green", alpha = .5) +
  geom_col(aes(y = new_cases), fill = "orange", alpha = .5) +
  scale_x_date(
    date_breaks = "1 week",
    date_labels = "%b %d",
    expand = c(0,.8)
  ) +
  scale_y_continuous(labels = scales::label_number(big.mark = ",")) +
  labs(
    title = "Number of New COV-19 Cases by Date",
    x = NULL, 
    y = "Newly Confirmed Cases", 
    caption = "Source: John Hopkins COV-19 Database"
  )
```



```{r, fig.width=5.5, fig.height=3.7}


# #New Cases by Date 
# carona_data_by_country_date %>% 
#   mutate(
#     new_cases = cases - lag(cases),
#     new_recovered = recovered - lag(recovered)
#   ) %>% 
#   ggplot(aes(x = date, y = new_cases, color = country_combined)) +
#   geom_point() +
#   geom_line() +
#   scale_x_date(
#     date_breaks = "1 week",
#     date_labels = "%b %d",
#     expand = c(0,.8)
#   ) +
#   scale_y_continuous(labels = scales::label_number(big.mark = ",")) +
#   theme(legend.position = "bottom") +
#   labs(
#     title = "Number of New COV-19 Cases by Country",
#     x = NULL, 
#     y = "Newly Confirmed Cases", 
#     color = NULL,
#     caption = "Source: John Hopkins COV-19 Database"
#   )
#add coord Cartesian # add the # of casese with str_glue 
```

### Country Information 

```{r, fig.width=5.5, fig.height=3.5}
#Countries Reporting Infections 
carona_data_usa %>% 
  filter(cases > 0) %>% 
  mutate(state = str_sub(state, start = 1, end = 3)) %>% 
  group_by(date) %>%
  summarise(num_states = n_distinct(state)) %>% 
  ggplot(aes(date, num_states)) +
  geom_col(fill = "orange") +
  scale_x_date(
    date_breaks = "1 week",
    date_labels = "%b %d",
    expand = c(0,.8)
  ) +
  labs(
    title = "Number of States Reporting Confirmed CoV cases",
    x = NULL, 
    y = "Number of States", 
    fill = NULL,
    caption = "Source: John Hopkins COV-19 Database"
  )

#Mortality rate by country 
carona_data_totals %>% 
  group_by(state) %>% 
  summarize(
    total_country_cases = sum(cases, na.rm = TRUE),
    total_country_deaths = sum(deaths, na.rm = TRUE),
  ) %>% 
  mutate(mortality_rate = total_country_deaths / total_country_cases) %>% 
  #filter(total_country_cases >= 20) %>% 
  arrange(desc(mortality_rate)) %>% 
  ggplot(aes(fct_reorder(state, mortality_rate, .desc = FALSE), mortality_rate)) + 
  geom_hline(
    aes(yintercept = sum(total_country_deaths)/sum(total_country_cases)),
    color = "red"
  ) +
  geom_col() +
  scale_y_continuous(labels = scales::label_percent()) +
  coord_flip() +
  labs(
    title = "Mortality Rate by State",
    y = "Crude Mortality Rate",
    x = NULL
  )
```

### Information and resources


CoV-19 is a is a novel Carona virus that was furst reported in Wuhan, China on
January 20, 2020. 

The virus is highly transmissable and spread:

- Between people who are in close contact with one another (within about 6 feet).

- Through respiratory droplets produced when an infected person coughs or sneezes.

People infeced with the virus are most contageous when they are sick, but 
in some cases the virus can be spread before symptoms arise

The [CDC Website has more useful information](https://www.cdc.gov/coronavirus/2019-ncov/about/transmission.html)

The [WHO Emergency Response site](https://experience.arcgis.com/experience/685d0ace521648f8a5beeeee1b9125cd)
also contains information on the global response to this epidemic. 

ESRI has created a similar dashboard that now requires login credentials to view
[link](https://coronavirus-disasterresponse.hub.arcgis.com/?adumkts=marketing&aduse=public_safety&aduc=industry_manager_outreach&utm_Source=industry_manager_outreach&aduca=cra_disaster_response_program&adut=cv-outreach-redirect&adulb=multiple&adusn=multiple&aduat=arcgis_online_portal&adupt=community&sf_id=701f2000000n6XKAAY)

All data for this dashboard was taken from the John's Hopkins daily CoV 
database hosted on [humanitarian data exchange]
(https://data.humdata.org/dataset/novel-coronavirus-2019-ncov-cases)


The 2020 COV outbreak is an ongoing public health crisis and this dashboard will
automatically update as more case data becomes availible

For any questions please contact: Christopher LeBoa (cleboa@stanford.edu)

**Things to add**
- mortality rate by country  -- done 
- add country name to table if > 50 cases 
- add # cases to str glue cumulative graph -- done
- sizing of leaflet dots -- ask bill 
- change C0v-19 by country to a line graph -- done
- # of new countries reporting cases per day -- done 


